### 2023-11-14 paper.tex

\documentclass[
  man,
  a4paper,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\makeatletter\makeatother\makeatletter\makeatother\makeatletter\@ifpackageloaded{caption}{}{\usepackage{caption}}\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}\makeatother\makeatletter\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}\makeatother\makeatletter\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}\makeatother\makeatletter\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}\makeatother\makeatletter\makeatother\makeatletter\makeatother

% \usepackage[style=apa,backend=biber]{biblatex}
% % \addbibresource{references.bib}
%

\title{Getting Started with the Graded Response Model (GRM): A gentle
introduction and tutorial in R}
\shorttitle{Graded Response Model Tutorial}

\authorsnames[{1},{1}]{
Ana Fulana,Blanca Zutana
}

\authorsaffiliations{
{Department of Psychology, Ludwig-Maximilians-Universität München,
Munich, Germany}}

\date{}
\abstract{Lorem ipsum dolor sit amet, consectetur adipisici elit, sed
eiusmod tempor incidunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquid
ex ea commodi consequat. Quis aute iure reprehenderit in voluptate velit
esse cillum dolore eu fugiat nulla pariatur. Excepteur sint obcaecat
cupiditat non proident, sunt in culpa qui officia deserunt mollit anim
id est laborum.}
% % \addbibresource{references.bib}
% 
\keywords{meta science, replication, multiverse analysis}

\authornote{\par{\addORCIDlink{Ana Fulana}{0000-0000-0000-0000}}
\par{ }
\par{This study was preregistered at OSF at \ldots. Code and data of
this study are available at \ldots{} This article is based on data
published in \ldots{} We declare we have no competing interests. This
research was funded by the German Research Foundation (DFG SCHO \ldots,
Blanca Zutana). We would like to thank Carina Mengana and Dolorita
Perengana for their very helpful comments and discussions on this
project. Because the authors are equal contributors, order of authorship
was determined by a fair coin
toss.  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: \emph{Ana
Fulana}: Conceptualization, Writing - original draft; \emph{Blanca
Zutana}: Project administration}
\par{Correspondence concerning this article should be addressed to Ana
Fulana, Email: sm@example.org}
}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
\hypertarget{quarto}{%
\subsection{Quarto}\label{quarto}}

Quarto enables you to weave together content and executable code into a
finished document. To learn more about Quarto see
\url{https://quarto.org}.

\hypertarget{running-code}{%
\subsection{Running Code}\label{running-code}}

When you click the \textbf{Render} button a document will be generated
that includes both content and the output of embedded code. You can
embed code like this:

\begin{verbatim}
[1] 2
\end{verbatim}

You can add options to executable code like this

\begin{verbatim}
[1] 4
\end{verbatim}

The \texttt{echo:\ false} option disables the printing of code (only
output is displayed).

\end{document}